---
title: "Assignment 11 - Regression Analysis"
author: "Suman K Polavarapu"
date: "November 8, 2015"
output: html_document
---

###Using R's lm function, perform regression analysis and measure the significance of the independent variables for the following two data sets:

```{r}
age   <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
maxHR <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 183, 178)
heartData <- data.frame(age, maxHR)
```

__Is the effect of Age on Max HR significant?__

$H_0$ : Age has no effect on Max HR ==>  $b_1 = 0$ , here  $b_1$ is the 'size of the effect' from the linear regression equation $y = b_0 + b_1 x + e$  \

$H_1$ : Age has effect on Max HR    ==>  $b_1 \neq 0$

Lets now use the R's built-in lm() function to generate the linear regression model to fit the above:

```{r}
fit.heartdata <- lm(maxHR ~ age, data = heartData) 
fit.heartdata
```

Hence, the linear model for the above data follows  : $$ MaxHR = 210.0485 - 0.7977Age $$

Now, lets see how significant the model is:

```{r}
(hdsummary <- summary(fit.heartdata))
```

From the above, we can reject the null hypothesis $H_0$ ( no effect on Max HR) due to the low significant level of the relationship, which
is much less than 0.01. (probability less than the significant threshold)

```{r}
round(hdsummary$coefficients["age", 4], 10) 
```

And we can conclude that there is correlation between Age and MaxHR.

```{r}
library(ggplot2)
ggplot(heartData, aes(age, maxHR)) + geom_point(aes(y=maxHR)) + stat_smooth(method = lm, level = .95) + xlab("Age") + ylab("Max Heart Rate") + ggtitle("Max Heart Rate Vs Age")
```

###Using the Auto data set perform a Linear Regression analysis using mpg as the dependent variable and the other 4 (displacement, horse-power, weight, acceleration) as independent variables.


$H_0$ : displacement, horse-power, weight, acceleration have ~NO~ effect on fuel efficiency,which means $b_1, b_2, b_3, b_4 = 0$ \

$H_1$ : displacement, horse-power, weight, acceleration have effect on fuel efficiency,which means $b_1, b_2, b_3, b_4 \neq 0$


```{r}
autodata <- read.table('auto-mpg.data',
                   col.names = c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg'))

head(autodata)
```


#### Perform linear regression using 40 random data rows from the data set.

Lets take random sample of 40 rows and calc the linear model, and measure the 95% confidence interval for each of the independent variable

```{r}
set.seed(10)
random40 <- autodata[sample(nrow(autodata), 40), ]
(auto.fit <- lm(mpg ~ . , data = random40))
summary(auto.fit)
auto.fit$coefficients
```


The equation is:
$$mpg = 44.12 - 0.02 Displacement - 0.006 Horsepower - 0.005 Weight + 0.076 Acceleration$$

Here, the low p-value suggests us to reject the null hypothesis ( the independent variables do not affect the dependent variable)

Lets measure the 95% conf interval
```{r}
confint(auto.fit, level=0.95)
```

Otherthan the intercept, none of the variables appears to be 'much' significant. Lets repeat the same on the entire data:

```{r}
(auto.fit.full = lm(mpg ~ . , data = autodata))
summary(auto.fit.full)
confint(auto.fit.full, level = 0.95)
```

Using the full data we get the linear fit as: 
$$ mpg = 45.25 - 0.006 Displacment - 0.043 Horsepower - 0.005 Weight - 0.023 Acceleration $$

And from the above data, in addition to the intercept, the Weight and Horsepower independent variables are showing significant impact on the fuel efficiency. Notice the negative impact they have on the auto mpg. [ The Weight's significant level is shown as $2.3 \times 10^{-10}$ and the 95% CI is -0.0036]. For other variables ( displacement, and acceleration , the significant level and the 95% CI do not rule out the null hypothesis).


```{r}
ggplot(autodata, aes(x=weight, y=mpg)) + geom_point() + stat_smooth(method = lm, level = .95) + xlab("Weight") + ylab("mpg") + ggtitle("Vehicle Weight vs Fuel Efficiency - Linear Regression 95% CI")
```

