---
title: "IS605 - Assignment 5"
author: "Suman K Polavarapu"
date: "Sunday, September 27, 2015"
output: html_document
---

### Problem Set 1
In this problem set we'll work out some properties of the least squares solution that we reviewed in the weekly readings. Consider the unsolvable system $Ax = b$ as given below:

$$
\begin{bmatrix} 1 & 0 \\ 1 & 1 \\ 1 & 3 \\ 1 & 4 \end{bmatrix}
\begin{bmatrix} x_1 \\ x_2 \end{bmatrix} =
\begin{bmatrix} 0 \\ 8 \\ 8 \\ 20 \end{bmatrix}
$$

#### Write R markdown script to compute $A^T A$ and $A^T b$

```{r}
#Prepare the matrix A, and the constraint vector b matrix.
A <- matrix(c(1,0,1,1,1,3,1,4), byrow=TRUE, nrow=4)
b <- matrix(c(0,8,8,20), nrow=4)

#Let's compute the $A^T A$ and $A^T b$
(ATA <- t(A) %*% A)
(ATb <- t(A) %*% b)
```

#### Solve for $\hat{x}$ in R using the above two computed matrices.

From the formula $A^T A \hat{x} = A^T b$ , the $\hat{x} = (A^T A)^{-1} (A^T b)$
```{r}
(x_hat <- solve(ATA) %*% ATb)

#Compare it with built-in R function 'lsfit' , which provides the least squares estimate of b, just like above.
x_hat_R <- lsfit(A, b, intercept=FALSE)
round(x_hat) == round(x_hat_R$coefficients)
```

#### What is the squared error of this solution?

The projection $p = \hat{x_1} a_1 + \hat{x_2} a_2$, where $\hat{x_1} and \hat{x_2}$ indicate 'estimates'
From the above, p can be written as , $p=A \hat{x}$. We already have the $\hat{x}$, lets find the projection vector p.
Also, We know that, $b = p + e$ , so the error vector $e = b - p$ , which is what we need to minimize :
```{r}
(p <- A %*% x_hat)

#Now , lets compute the error vector e , which is b - p
(e <- b - p)

# Compare the error vector, with built-in lsfit function - residuals.
round(e) == round(x_hat_R$residuals)

#Find the squared error
(sq.e <- sum(e^2))

```

#### Instead of b = [0; 8; 8; 20], start with p = [1; 5; 13; 17] and find the exact solution (i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector e = 0).

```{r}
#basically, the case where b=p.

#Let's first find the A^T p , we have already got the (A^T A)A -1.
p = matrix(c(1,5,13,17), nrow=4)
(ATp <- t(A) %*% p)
```

_Find x-hat, by finding $(A^T A)^{-1} (A^T p)$_

```{r}
(x_hat_p <- solve(ATA) %*% ATp)

#p=Ax, and then, calculate the error
(p2 <- A %*% x_hat_p)
(error <- round(p2 - p))
(sq.error <- round(sum(error^2), digits=6))
```
So, from the above, we have solvable system of equations.

#### Show that the error e = b - p = [-1; 3;-5; 3].

```{r}
(res.error <- b - p)
```

#### Show that the error e is orthogonal to p and to each of the columns of A.

If 2 vectors are orthogonal , then their dot product is equal to zero. 

```{r}

#e and p are orthogonal
round(sum(e*p))

#e and the columns of A are orthogonal
round(sum(e*A[,1]))
round(sum(e*A[,2]))
```


### Problem Set 2
Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learning dataset). This dataset contains 5 columns: displacement, horsepower, weight, acceleration, mpg. We are going to model mpg as a function of the other four variables.

```{r}
data <- read.table('auto-mpg.data', 
                       col.names=c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg'))

summary(data)
```

Lets convert the data frame into $A$ and $b$ matrices

```{r}
A <- data.matrix(data[,1:4])
A <- cbind(rep(1, nrow(A)), A)  #added a column of 1's for the intercepts
b <- data.matrix(data[,5])
head(A)
head(b)
```

We got the required inputs, to calculate the coordinates $\hat{x}$ which approximates the solution to the $Ax = b$.

```{r}
ATA <- t(A) %*% A
ATb <- t(A) %*% b

#Lets calculate the x_hat
(x_hat <- solve(ATA) %*% ATb)
```

The above $\hat{x}$ gives us the best fitting solution for the given auto mpg data, Lets find out the error.

```{r}
p <- A %*% x_hat
e <- b - p
(sq_e <- sum(e^2))
```

Let's match with R built-in function lsfit
```{r}
lsr <- lsfit(A, b, intercept=FALSE)  #we already added 1st vector for intercept, so no need to add again.
lsr$coefficients

round(x_hat) == round(lsr$coefficients)
```