---
title: "IS605 - Assignment 1"
author: "Suman K Polavarapu"
date: "Saturday, August 29, 2015"
output: html_document
---

## PROBLEM SET 1 :


__1. Calculate the dot product u.v where u = [0.5, 0.5] and v = [3, -4]__

```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)
(dot.product <- u %*% v) #This returns the inner product as a matrix

#Alternatively, We can simply use the sum(u*v)
(dot.product <- sum(u*v))
```


__2. What are the lengths of u and v ?__

```{r}
( u.length <- sqrt( sum(u * u) ) )
( v.length <- sqrt( sum(v * v) ) )
```


__3. What is the linear combination: 3u - 2v ?__

```{r}
( linearComb <- 3 * u - 2 * v )
```


__4. What is the angle between u and v ? __

The dot product of the vectors u and v is defined as:

u . v = ||u||  ||v||  cos θ

So, θ = arccos ( u . v / ||u||  ||v|| )

```{r}
#Angle between u and v vectors
(angle.radians <- acos( sum(u * v) / ( sqrt(sum( u * u)) * sqrt(sum(v * v)) )) ) 
#In Degrees
(degrees <- angle.radians * (180/pi))

````


## PROBLEM SET 2 :

__1. Function to solve the system of equations with 3 variables and 3 constraints, solving for x__

```{r}

#Solves 3  X 3 system of equations using the below :
#systematically convert the co-efficient matrix A into the Upper Triangular Form as below:
#(1) Start with row 1 of the co-efficient matrix
#(2) Pivot: The first non-zero element in the row being evaluated
#(3) Multiplier: The element being eliminated divided by the Pivot
#(4) Subtract Multiplier times row n from row n+1
#(5) Advance to the next row and repeat

#Inputs  - Co-efficient matrix, and the constrains vector.
#Output  - a vector x of solutions for x1,x2 and x3.

solve3by3System <- function(A,b)
{

  #creat a numeric vector of size 3, to keep outputs.
  x = numeric(3);
  
  #column bind the constraints vector.
  (A<-cbind(A,b))
  
  #start with row1, check the pivot, if zero swap.
  if(A[1,1] == 0)
  {
    if (A[2,1] != 0) 
    {
      A <- A[c(2,1,3),]
    }
    else
    {
      A <- A[c(3,1,2),]
    }
  }

  #row2 elimination - subtract the [multiplier * row1] from row2, by using A[1,1] as pivot
  multiplier <- (A[2,1]/A[1,1])
  A[2,] <- A[2,] - (multiplier * A[1,])
  
  #row3 elimination - subtract the [multiplier * row1] from row3
  multiplier <- (A[3,1]/A[1,1])
  A[3,] <- A[3,] - (multiplier * A[1,])
  
  #swap for zero pivot - if A[2,2] is zero then swap with row 3
  if (A[2,2] == 0)
  {
    A <- A[c(1,3,2) ,]
  }
  
  #row3 elimination - Now, use the A[2,2] as the pivot for row3 elimination
  multiplier <- (A[3,2]/A[2,2])
  A[3,] <- A[3,] - (multiplier * A[2,])
  
  #Back substitution step
  x[3] <- A[3,4]/A[3,3]
  x[2] <- (A[2,4]-A[2,3]*x[3])/A[2,2]
  x[1] <- (A[1,4]-A[1,2]*x[2] - A[1,3]*x[3])/A[1,1]
  
  return (round(x,2))
}
```

__Client:__
```{r}
#Inputs...
(A<- matrix(c(1,1,3,2,-1,5,-1,-2,4), ncol=3, byrow=T))
(b<- matrix(c(1,2,6)))
#client call...
(solve3by3System(A,b))
```

__Trying the same with x <- (A Inverse) X b , using built-in inverse method in R, "solve"__

Ax = b ==> x = Inverse-of-A * b, thus x can be derived by inversing A and mulitiplying with the constraint vector b.
```{r}

(x <- round( solve(A) %*% b , 2 ))

```
