---
title: "IS605 - Assignment 2"
author: "Suman K Polavarapu"
date: "Saturday, September 05, 2015"
output: pdf_document
---

## Problem Set 1 :

### (1) Show that \(A^TA \neq AA^T\) in general.

Let \(A\) be an m x n matrix , so \(A^T\) is an n x m matrix. By multiplying A by \(A^T\) produces an m x m matrix.
And multiplying \(A^T\) with A would produce n x n matrix. So, the matrices would clearly differ in dimension, when m != n. 

Lets check a case where m == n. 

\(A=\begin{bmatrix}
   a & b & c \\
   d & e & f \\
   g & h & i
\end{bmatrix}\)
  
\(A^T=\begin{bmatrix}
   a & d & g \\
   b & e & h \\
   c & f & i
\end{bmatrix}\)

\(AA^T=\begin{bmatrix}
  a^2+b^2+c^2 & ad+be+cf & ag+bh+ci \\
  \cdots      & \cdots   & \cdots   \\
  \cdots      & \cdots   & \cdots 
\end{bmatrix}\)

However, \(A^TA=\begin{bmatrix}
  a^2+d^2+g^2 & ab+de+gh & ac+df+gi \\
  \cdots      & \cdots   & \cdots   \\
  \cdots      & \cdots   & \cdots 
\end{bmatrix}\)

Notice the mismatch of the dot products of the first row elements itself. So, generally \(A^TA \neq AA^T\)


### (2) For a special type of square matrix A, we get \(A^TA = AA^T\) . Under what conditions could this be true?

\(A^TA = AA^T\) is true , if a matrix is a square matrix and all of its elements are same, Or, if the matrix is an identity matrix 
(/or a scalar multiples of it).

```{r}
A <- matrix(c(5,5,5,5,5,5,5,5,5), nrow=3, byrow=T)
A  
t(A)%*%A == A%*%t(A)
```
  
  
Similarly, a diagonal matrix. 
```{r}
A <- matrix(c(5,0,0,0,5,0,0,0,5), nrow=3, byrow=T)
A  
t(A)%*%A == A%*%t(A)
```


## Problem Set 2 :

### Write an R function to factorize a sqaure matrix A into LU or LDU.

```{r}
###################
#Decompose/Factorize a given matrix into L and U.
#Inputs: A - matrix, which need to be factorized. 
#        We will apply the elimination steps to this, which eventually be our U.
#Output: A list of L and U.
###################

FactorizeLU <- function(U) {
  
  rows = nrow(U)
  cols = ncol(U)
  
  if ( rows != cols) {
    print("*** Given Matrix is NOT a square matrix !!! *** ")
  }
  
  #Identity matrix, to store the multipliers used in elimination, this would be 
  # eventually our lower triangular matrix.
  L <- diag(rows)
  
  #Start eliminating the E21, (/E31, E32 etc..) from the matrix U, 
  #and in that process capture the E (elimination steps) as well
  for(i in 2:rows) {
    
    for(j in 1:(i-1)) {
      
      #Generate Elimination matrix. 
      #For row2 , it would be E21, For row3, it would be E31, E32 etc..
      E <- diag(rows)
      E[i, j] <- -(U[i,j]/U[j,j])  #-(multiplier), for E21, it would be - U[2,1]/U[1,1]
      
      #Eliminate the elements from Upper triangular matrix.Because,  U = (E32E31E21) A 
      U <- E %*% U
      
      #keep the elimination step in L. (the multiplier used in the elimination)
      #because, L <- solve(E21) %*% solve(E31) %*% solve(E32)
      L <- L %*% solve(E)
      
    } # for each column
  } # for each row
    
  return(list('L'=L,'U'=U))  #return L as L , and U as U attribute.
  
}
```

#### Client calls:
```{r}
(A <- matrix(c(2,1,6,5), nrow=2))
(res <- FactorizeLU(A))
(res$L %*% res$U == A)
````

```{r}
(A <- matrix(c(1, 2, 1, 3, 4, 1, 5, 7, 0), nrow=3))
(res <- FactorizeLU(A))
(res$L %*% res$U == A)
```


### LU Decomposition using package - matrixcalc

```{r}
library(matrixcalc)
A <- matrix( c (1, 2, 1, 3, 4, 1, 5, 7, 0 ), nrow=3)
luA <- lu.decomposition( A )
(L <- luA$L)
(U <- luA$U)
(luA$L %*% luA$U == A)
```