---
title: "Assignment 9"
author: "Suman K Polavarapu"
date: "October 25, 2015"
output: html_document
---

### Problem Set

#### write a function that will produce a sample of random variable that is distributed as follows:

$$
f(x) = \left\{ \begin{array}{1 1}
x \quad 0 \leq x < 1 \\
2-x \quad 1 \leq x \leq 2
\end {array} \right.
$$


__This function can be simply defined as below, which can be later used in R's built in 'sample' method to generate samples__

```{r}
fun1 <- function(x){
    if(0 <= x && x <= 2){
        if(x <= 1){
            return(x)
        } else {
            return(2-x)
        }
    }
}
```


__However, let's use the Inverse Sampling Technique here, to generate samples :__

Create an inverse cdf function for this distribution:

Anti-derivative for the above function is:

$$
F(x) = \left\{ \begin{array}{1 1}
\frac{x^2}{2} \quad 0 \leq x < 1 \\
2x - \frac{x^2}{2} \quad 1 \leq x \leq 2
\end{array} \right.
$$

The inverse of the anti-derivative:
$$
F^{-1}(x) = \left\{ \begin{array}{1 1}
\sqrt{2y} \quad 0 \leq y < \frac{1}{2} \\
2 - \sqrt{2(1 - y)} \quad \frac{1}{2} \leq y \leq 1
\end{array} \right.
$$

__The Inverse CDF function:__

```{r}
inverseCDF <- function(y){
  resp <- 0
  if (0 <= y &&  y <= 1) {
    if ( y <= 0.5) {
      resp = (sqrt(2*y))
    } else {
      resp = (2 - sqrt(2 * (1-y)) )
    }
  }
  else  {
    resp = 0 # Out of range !
  }
  return (resp)
}

```

Lets apply the above inverse cdf function to a vector of randomly selected values between 0 and 1, and plot those data points.

```{r}
require(ggplot2)
sample.data <- data.frame(X = sapply(runif(10000), inverseCDF))
print(head(sample.data))
ggplot(data = sample.data, aes(sample.data$X)) + geom_histogram(binwidth=.05) + ggtitle("Triangular PDF")
```


####Now, write a function (using R) that will produce a sample of random variable that is distributed as follows:

$$
f(x) = \left\{ \begin{array}{1 1}
1-x \quad 0 \leq x \leq 1 \\
x-1 \quad 1 < x \leq 2
\end {array} \right.
$$

```{r}
fun2 <- function(x) {
  resp <- 0
  if (0 <= x && x <= 2) {
    if (x <= 1) {
      resp <- (1 - x)
    }
    else {
      resp <- (x - 1)
    }
  }
  
  return (resp)
}
```

First create a vector of probabilities using the above function.And use that as 'prob' for R-built in function 'sample', and plot.
```{r}
#simply, create vector of values between 0 and 2
x <- seq(from = 0, to = 2, length.out = 10000)

#create vector of probabilities, which can be used later for sampling.
funx.prob <- sapply(x, fun2)

#Now, create a sample
sample.df <- data.frame(X = sample(x, 10000, replace=TRUE, prob=funx.prob))
ggplot(sample.df, aes(sample.df$X)) + geom_histogram(binwidth=.05) + ggtitle("Inverted Traingular PDF")

```


####Write a program that will take a sample set size (n) as a parameter and the PDF as a second parameter,and perform 1000 iterations where it samples from the PDF, each time taking n samples and computes the mean of these n samples. It then plots a histogram of these 1000 means that it computes.

```{r}
distributionOfMeans <- function(n, pdf) {
   #Calculates the mean of 1000 samples of size n for the given pdf
   #plots the distribution of the mean.
  
  x <- seq(0, 2, by=0.01)
  funx.prob <- sapply(x, pdf)
  means <- c()
  
  for(i in 1:1000) {
    sampleiter <-  sample(x, n, replace=TRUE, prob=funx.prob)
    means <- c(means, mean(sampleiter))
  }
  
  print(paste(("The mean of the samplings is: "), mean(means)))
  print(paste(("The standard deviation of the samplings is: "), sd(means)))
  
  means.df <- data.frame(means)
  ggplot(means.df, aes(means)) + geom_histogram(binwidth=.025) + xlab("Mean of the Sample") + ggtitle(paste("Distribution of Means;PDF=",  deparse(substitute(pdf)), ";Samples=1000;Sample Size=", n)) 
  

}

```

__Client Calls:__

Lets try with the sample size of 20
```{r}
distributionOfMeans(n=20, pdf=fun1)
distributionOfMeans(n=20, pdf=fun2)
```

Lets try with the sample size of 10
```{r}
distributionOfMeans(n=10, pdf=fun1)
distributionOfMeans(n=10, pdf=fun2)
```

From the above, its evident that even for reasonably small sample sizes such as 10, Central Limit Theorem holds true - The means of the sampling distributions from 'any' probability distribution, forms a bell shaped curve ('normally distributed' around the mean of the original distribution)